---
title: "EL NINO (UCI/ML)"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r, echo=TRUE, warning=FALSE, message=FALSE}
library(ggplot2) # Data Visualisation
library(moments) # Moments: skewness, kurtosis
```


### Importing table and assigning column
```{r, cache=TRUE}
taoall <- data.table::fread('elnino/tao-all2.dat.gz')

colnames(taoall) <- c("obs", "year", "month","day", "date", "lat", "lng", 
                      "zon_wind", # Zonal wind - for west, + for east
                      "mer_wind", # Meridional wind - for south, + for north
                      "humidity", "air_temp", "sub_temp"
                      )

```

### Column conversion and Basic stats
```{r, warning=FALSE}
## Conversion of the variables into numeric (right format for columns)
taoall$zon_wind <- as.numeric(taoall$zon_wind)  
taoall$mer_wind <- as.numeric(taoall$mer_wind)
taoall$humidity <- as.numeric(taoall$humidity)
taoall$air_temp <- as.numeric(taoall$air_temp)
taoall$sub_temp <- as.numeric(taoall$sub_temp)

taoall$date <- lubridate::ymd(taoall$date)  # Formatting data column

### nrow(taoall)
### Summary for data set
summary(taoall)
```

### Some more basic Stats
```{r}
###  Standard deviation 
apply(na.omit(taoall[, 8:12]), 2, sd)

### Inter Quartile Range
apply(na.omit(taoall[, 8:12]), 2, IQR)

### Skew
apply(na.omit(taoall[, 8:12]),2, skewness)

### kurtosis
apply(na.omit(taoall[, 8:12]),2, kurtosis)
```
### Na's Exploration
```{r}  
## Na's Exploration
a <- sum(is.na(taoall$zon_wind))/nrow(taoall) # NA's in percent for zonal wind
b <- sum(is.na(taoall$mer_wind))/nrow(taoall) # NA's in percent for meridional wind
c <- sum(is.na(taoall$humidity))/nrow(taoall) # NA's in percent for humidity
d <- sum(is.na(taoall$air_temp))/nrow(taoall) # NA's in percent for air temprature
e <- sum(is.na(taoall$sub_temp))/nrow(taoall) # NA's in percent for subsurface temprature


### Printing Data Frame

(data.frame(Variable = c("Zonal Wind", "Meridional Wind", "Humidity", "Air Temprature", "Subsurafe Temprature"),
            Sum = c(sum(is.na(taoall$zon_wind)),sum(is.na(taoall$mer_wind)),sum(is.na(taoall$humidity)),sum(is.na(taoall$air_temp)),
                    sum(is.na(taoall$sub_temp))),
            Percentage = c(a, b, c, d ,e)
              ))


```
### Univariate Analysis
```{r, cache=TRUE}
############### Univariate Exploration ########################
par(mfrow = c(2,3))
hist(taoall$zon_wind);abline(v = mean(na.omit(taoall$zon_wind)), col = 'red')
hist(taoall$mer_wind);abline(v = mean(na.omit(taoall$mer_wind)), col = 'red')
hist(taoall$humidity);abline(v = mean(na.omit(taoall$humidity)), col = 'red')
hist(taoall$air_temp);abline(v = mean(na.omit(taoall$air_temp)), col = 'red')
hist(taoall$sub_temp);abline(v = mean(na.omit(taoall$sub_temp)), col = 'red')
```

### Normality Visualisation
```{r, cache=TRUE}
### Normality check
par(mfrow = c(2,3))
qqnorm(taoall$zon_wind, main = "Zonal Wind"); qqline(taoall$zon_wind)
qqnorm(taoall$mer_wind, main = "Meridional Wind"); qqline(taoall$mer_wind)
qqnorm(taoall$humidity, main = "Humidity"); qqline(taoall$humidity)
qqnorm(taoall$air_temp, main = "Air Temprature"); qqline(taoall$air_temp)
qqnorm(taoall$sub_temp, main = "Subsurface Temprature"); qqline(taoall$sub_temp)
```



### Advance Plot
```{r, cache=TRUE, warning=FALSE, message=FALSE}
################ Visualization ###############################


p1 <- ggplot(na.omit(taoall), aes(date, zon_wind))+
  geom_hex()+
  geom_smooth(method = "lm")+
  geom_smooth()+
  theme_classic()

p2 <- ggplot(taoall, aes(date, mer_wind))+
  geom_hex()+
  geom_smooth(method = "lm")+
  geom_smooth()+
  theme_classic()

p3 <- ggplot(taoall, aes(date, air_temp))+
  geom_hex()+
  geom_smooth(method = "lm")+
  geom_smooth()+
  theme_classic()


p4 <- ggplot(taoall, aes(date, sub_temp))+
  geom_bin2d( )+
  geom_smooth(method = "lm")+
  geom_smooth()+
  theme_classic()
###################### Air ######################################
p5 <-  ggplot(taoall, aes(year, air_temp))+
  geom_hex()+
  geom_smooth(method = "lm")+
  geom_smooth()+
  theme_classic()
p6 <-  ggplot(taoall, aes(month, air_temp))+
  geom_hex()+
  geom_smooth(method = "lm")+
  geom_smooth()+
  theme_classic()

p7 <-  ggplot(taoall, aes(day, air_temp))+
  geom_hex()+
  geom_smooth(method = "lm")+
  geom_smooth()+
  theme_classic()

###################### Sub Surface Air ###########################


p8 <-  ggplot(taoall, aes(year, sub_temp))+
  geom_hex()+
  geom_smooth(method = "lm")+
  geom_smooth()+
  theme_classic()



p9 <-  ggplot(taoall, aes(month, sub_temp))+
  geom_hex()+
  geom_smooth(method = "lm")+
  geom_smooth()+
  theme_classic()

p10 <-  ggplot(taoall, aes(day, air_temp))+
  geom_hex()+
  geom_smooth(method = "lm")+
  geom_smooth()+
  theme_classic()

gridExtra::grid.arrange(p1, p2, p3, p4, p5, p6, p7, p8 , p9 , p10)
```


```{r}
par(mfrow = c(3,2))
ls <- apply(na.omit(taoall[,8:12]), 2, acf)
#par(mfrow = c(2,2))
#ls[1]; ls[2]; ls[3]; ls[4]

```
### Discussion 

This data set is downloaded from <http://archive.ics.uci.edu/ml/datasets/El+Nino>
machine learning repositories.Primarily used by Dr. Di Cook, Department of Statistics, Lowa State 
University, data set collected by the international Tropical Ocean Global Atmosphere
(TOGA) program using Tropical Atmosphere Ocean (TAO) array. Data set consist of 
11 variables and 178080 observations, of which year, month and day are just by part
of date column. Technically variable of interest in this case are  zonal wind, 
meridional wind, humidity, air temperature and sub surface temperature. And 
obviously time phenomena makes it more interesting and complex to study. Though the
variables of interest have unknown values in the data set, there is no technical 
explanation of which why this has happened. 

Data has observation from 1980,Aug 03 to 1998,June 23 for daily interval. Zonal wind has
minimum -12.4 and maximum 14.3, where (+) for east and (-) for west. Then meridional winds 
has minimum of -11.60 and maximum of 13.0, where (-) stands for south and (+) for north.
On the other hand, temperature for air and sub surface seem to be quit identical in basic
exploration have (17.05, 17.35) minimum, (31.66, 31.26) maximum and mean (26.89, 27.71) 
respectively. Even kurtosis for temperature has quite large and identical value of 5.44
and 5.77, which is excess kurtosis cases(>3). This data set is suffering from unknown data 
cells, which is quite high for humidity (36 percent). Clearly omitting these observation 
from the data set will cause lost of many information from the data. but a systematic 
analysis can give valuable insights. Meridional wind and humidity are seem to be normal
in initial graphical analysis. when zonal wind is left skewed and air temperature and 
sub suraface temprature are again characterizing identical right skewed. Results for 
normality is being anticipated by normality graphs, Meridional wind and Humidity is 
still seem to be normal while other are seem following normal for cluster of observation
around mean while tails are thick (non normal). From the advance analysis graph, trend for zonal 
wind seem to be stagnant over the period of time. Meridional wind is showing downward trend over
the years but stagnant between months and days. In contrast to this, air and sub surface temperature
are identical here too, increasing trend for years and constant for months and days. This is proved 
temperature for air and sub suraface air are highly correlated.



```{r}
corrplot::corrplot(cor(na.omit(taoall[,8:12])),type = "lower") 
```






























