---
title: 'Glass Data (UCI/ML) '
output:
  pdf_document: default
  html_document:
    df_print: paged
---


# Basic Statistics in Exploration
```{r}

glass <- data.table::fread("glass/glass.data")
colnames(glass) <- c("id","Ri","Na","Mg","Al","Si","K","Ca","Ba","Fe","type")

glass <- glass[,-1]
glass$type <- as.factor(glass$type)
  
summary(glass)
```

```{r}
library(moments)
# Inter-Quartile Range
apply(glass[,-"type"], 2, IQR)
# Standard deviation
apply(glass[,-'type'], 2, sd)

# skewness
apply(glass[,-"type"], 2, skewness)
 
# kurtosis
apply(glass[,-'type'],2, kurtosis)

```


# Univariate Graphical Analysis
```{r}
### Dependent (Type Variable)
names <- c("BW-Float", "BW-Non_Float", "VH-Float",  "Container", "Tableware", "Headlamps")
plot(glass$type, col = "black",
     ylim= c(0,90),
     # yaxt = "n",
     xlab = "Type of Glass",
     names = names)
```


```{r}
### Explanatory Variables
par(mfrow = c(3,3))
hist(glass$Ri);abline(v = mean(glass$Ri), col = "red")
hist(glass$Na);abline(v = mean(glass$Na), col = "red")
hist(glass$Mg);abline(v = mean(glass$Mg), col = "red")
hist(glass$Al);abline(v = mean(glass$Al), col = "red")
hist(glass$Si);abline(v = mean(glass$Si), col = "red")
hist(glass$K);abline(v = mean(glass$K), col = "red")
hist(glass$Ca);abline(v = mean(glass$Ca), col = "red")
hist(glass$Ba);abline(v = mean(glass$Ba), col = "red")
hist(glass$Fe);abline(v = mean(glass$Fe), col = "red")
```


```{r}
corrplot::corrplot(cor(glass[,-"type"]),type = "upper", method = "number")

```
### Diagonostic 
```{r}
par(mfrow = c(3,3))
qqnorm(glass$Ri);qqline(glass$Ri)
qqnorm(glass$Na);qqline(glass$Na)
qqnorm(glass$Mg);qqline(glass$Mg)
qqnorm(glass$Al);qqline(glass$Al)
qqnorm(glass$Si);qqline(glass$Si)
qqnorm(glass$K);qqline(glass$K)
qqnorm(glass$Ca);qqline(glass$Ca)
qqnorm(glass$Ba);qqline(glass$Ba)
qqnorm(glass$Fe);qqline(glass$Fe)
```

## Multivariate
```{r}
pairs(glass)
```

### Ri vs most correlated variables (Can be use for Modelling)
```{r}
par(mfrow = c(1,2))
plot(Ri~Ca, glass); abline(lm(Ri~Ca, glass))
plot(Ri~Si, glass); abline(lm(Ri~Si, glass))
```


```{r}
## Normality Test
apply(glass[,-"type"], 2, shapiro.test)
```



## Type vs Independents 
```{r}
### Dependant to independent
par(mfrow = c(3,3))


boxplot(Ri~type, data = glass,
        col = topo.colors(6))

boxplot(Na~type, data = glass,
        col = topo.colors(6))

boxplot(Mg~type, data = glass,
        col = topo.colors(6))

boxplot(Al~type, data = glass,
        col = topo.colors(6))

boxplot(Si~type, data = glass,
        col = topo.colors(6))

boxplot(K~type, data = glass,
        col = topo.colors(6))

boxplot(Ca~type, data = glass,
        col = topo.colors(6))

boxplot(Ba~type, data = glass,
        col = topo.colors(6))

boxplot(Fe~type, data = glass,
        col = topo.colors(6))

```

### Description and Findings

Purpose of this analysis is to analyze descriptively for relations,  associations among
variables and individual characteristics variables. Glass Identification data set has taken
from<https://archive.ics.uci.edu/ml/datasets/glass+identification>, originally
came from Vina Spiehler, Ph.D., DABFT Diagnostic Products Corporations. In data
set, we have total 9 variables and 214 observations to study. Moving to analysis
part, here we have two objective variable which are Refractive index and Type of 
glass to study. Refractive index is a continuous attribute on the other hand type 
of glass has seven classes (class 4 is missing in data). In the univariate graphical
section, we have individual graphs of each attribute. For most of attributes 
cluster of observations is around its mean in accompany with its mode e.g. mean of
Na (Sodium) is 13.41 and the median is 13.30. Interesting characteristics is came into
light for attributes K and Ba, majority of observations are extreme e.g. for K(Potassium)
between 0 to 1 and for Fe (Iron) its zero. Interestingly Mg (Magnesium) has two peaks
(bi-modal) at both extremes. All these attributes are sharing asymmetric characteristics so
they are likely to be non-normal. For type variable, there are 6 classes present in the
table where Building window floating and non floating class has dominance, on the other 
hand Tableware and Container are at bottom in numbers for data set (13 and 9respectively).

For association between variables, I have drawn correlation matrix which depicts variables 
, which are highly correlated with Iron, If iron increases by 1 unit so Ca will increase 
by .8 and Si (silicon) will fall for .5.Most highly correlated variables can be used for modelling
Correlation does not implies causality. It only tells the direction of change in variables. Normality is seen
normal for most variables but their tails are thick (except Mg, Ba, K and Fe).Since the p
value is less than .05, Normality for all the attributes is rejected by Shapiro Wilkson 
test. 
For part b) I have drawn box-plot, also called box-whisker, which uses to show categorical 
variation of a variable against another continuous variable. Median for Ri and Fe are quite
same across all the type of glasses. Magnesium has higher value for building windows of 
either type and vehicle window float processed and higher spread for tableware and
container.

